### 2008-09-02 draft/ConFut.tex
First add of the master draft

\chapter{Discusion and Conclusion}

\section{  Conclusion}
\label{sec:ConclusionConclusion }

\section{Future Work}
\label{sec:FutureWork}