### 2008-09-02 draft/System.tex

\chapter{Proposed System for Sketch Understanding}
\section{Aims of the System}
\label{sec:AimsOfTheSystem}
The previous chapter shows the current trends in sketch recognition. This thesis introduce a solution for the problem that give optimize solution for the segmentation problem.

\section{An Overview of the System}
\label{sec:AnOverviewOfTheSystem}

\section{System Components}
\label{sec:SystemComponents}
% the system and discription of this

\chapter{Preprocessing and Segmentation}
\label{sec:Preprocessingandsegmentation}

\section{Preprocessing}
\label{sec:Preprocessing}

\subsection{Curvature Calculation}
\label{sec:CurvatureCalculation}

\subsection{Critical Point Detection}
\label{sec:CriticalPointDetection}

\section{Segmentation}
\label{sec:Segmentation}
\subsection{Swarm Introduction}
\label{sec:SwarmIntroduction}

\subsection{Ellipse Detection }
\label{sec:EllipseDetection}

\subsection{Swarm Segmentation}
\label{sec:SwarmSegmentation}
\subsubsection{Problem Formulation }
\label{sec:ProblemFormulation}

\subsubsection{Fitness Computation }
\label{sec:FitnessComputation}
\subsubsection{Polygon Division Algorithm}
\label{sec:PolygonDivisionAlgorithm}

\subsubsection{Curve Divisions Algorithm}
\label{sec:PolygonDivisionAlgorithm}

\section{BenchMarck algorithm}
\label{sec:BenchMarckAlgorithm}

\chapter{Symbol Recognition}
\label{sec:SymbolRecognition}
\section{Stroke Clustring Algoirthm }
\label{sec:ClustringAlgoirthm}
\section{ Feature Extraction}%%Feature Set
\label{sec:FeatureExtraction}

%%%%%%%%%%%%%%%%%%%55555555
%Symbol Recognition 16
%4.1 Preprocessing and segmentation . . . . . . . . . . . . . . . . . . . . . . . . . . . 16
%4.2 Feature-Based Statistical Symbol Recognition . . . . . . . . . . . . . . . . . . . . 18
%4.2.1 Feature Set . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 19
%4.2.2 Training . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 21
%4.2.3 Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 23
%4.2.4 User Study . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 24
%4.3 Graph-Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 28
%4.3.1 Previous Work . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 28
%4.3.2 Statistical Graph Definitions and Error-driven Stochastic Graph Mathcing . 29
%4.4 Image Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 30
%4.4.1 User Study . . . . . . . . . . . . . .
%3.3 System Components 43
%4 Segmenting a Freehand Sketch
%4.1 Interface 51
%4.2 Spline representation 52
%4.3 Points of potential feature separation 59
%4.4 Corner detection 59
%4.5 Intersection detection 61
%4.6 Clustering of interesting points 65
%4.7 Feature Extraction