### 2009-04-27 draft/SketchMaster.tex

%==============================================================================
%
%      File:  Master.TEX  (Stored in TEX$LATEX: as UNHTHESIS.TEMPLATE)
%  Language:  LaTeX - Document Preparation System
%      Date:  26/Dec/2005
%    Author:  Maha El Meseery
%             
%
%   This document  is my master thesis on sketch Recognition and understanding
%
%
%        % latex mythesis
%
%------------------------------------------------------------------------------

%    14 Aug 02 Updated document to latex 2e. Also, changed wording 
%              to be consistent with an X windows system.

%    14 Aug 02 Added \includeonly command in comments.

% During the editing process you will not want to view and print the
% whole document. Uncomment the command below so that only the
% chapters that you want to work on are visible.
%\includeonly{ChapterIntroduction}
%\documentstyle[11pt,unhthesis]{report}
\documentclass[11pt,doublespace]{Sketchthesis}

% Extra modules can be added to your document to extend the things
% the Latex can Do. One of these is psfig which allows you to print
% postscript files. You tell Latex to read these packages using the
% \usepackage command:

\usepackage{psfig}
\usepackage{named}
\usepackage{natbib}
\bibliographystyle{plain}
\hyphenation{gno-mon-ly}

\usepackage {graphicx}
\usepackage{subfigure}
\usepackage{longtable}
\usepackage{array}
%\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}

%------------------------------------------------------------------------------
%  Preliminary Pages - Fill in the `blanks' noted.
%------------------------------------------------------------------------------

\begin{document}
                                                        % TITLE PAGE
                                                        %======================
\title{Sketch Recognition and Understanding}             % My theses title  using Particle Swarm Optimization Algorithm
\author{Maha Mohammed Nabeel El Meseery}                               % Y name
\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
\major{Computer Engineering}                            % My current major
\degree{Master of Science}                              % My new degree
\degreemonth{}                                      % When awarded. to be added when known
\degreeyear{}                                      %
%th today or first day of writing  December 26 2005 
%\thesisdate[]                                     % Date of document.
\DOCUMENTtype{Thesis}                                   % or DISSERTATION
\Documenttype{Thesis}                                   % or Dissertation
\documenttype{Thesis}                                   % or dissertation
\maketitle

\providecommand{\tabularnewline}{\\}
                                                        % COPYRIGHT PAGE
                                                        %======================
%\copyrightyear{2009}                                    % Delete these
%\makecopyright                                          % if no copyright
                                                    % APPROVAL                                                         %======================
%\thispagestyle{empty}               % Don't do headers or footers.
%\newpage                                                             
%\begin{center}     
%\LARGE  
%\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
%\normalsize by      \\ 
%\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
%\vspace*{0.4in}        
%\normalsize
%A Thesis Submitted To the\\
%Faculty of Engineering, Cairo University\\
%in Partial Fulfillment of the\\
%Requirements for the Degree of\\
%MASTER OF SCIENCE
%in\\
%COMPUTER ENGINEERING\\
%
%\end{center}
%          
%    Approved by\\                                           
%Examining Committee:\\
% \makebox[6in]{\hrulefill}\\ 
%Professor Nevin Darwish, Thesis Main Advisor\\
% \makebox[6in]{\hrulefill}\\ 
%Professor Samia Mashali,Thesis Advisor \\
% \makebox[6in]{\hrulefill}\\  
%Assoc. Professor Magda Fayek,Thesis Advisor\\ 
% \makebox[6in]{\hrulefill}\\ 
%Professor  Ashraf Abd El Wahab, Member\\
% \makebox[6in]{\hrulefill}\\ 
%Professor Mohamed Zaki Abd ElMegeed, Member\\
% \makebox[6in]{\hrulefill}\\ 
  % \\   you need...\\
%\makeapproval                                           %

                                                        % DEDICATION PAGE
                                                        %======================
%\begin{dedication}                                      % Delete these
												                                  % if no dedication
%\end{dedication}                                        % page.

                                                        % ACK. PAGE
                                                        %======================
\begin{acknowledgments}                                 % Delete these if
All praise is due to Allah Who guided me to this. I could not truly have been led aright if
Allah had not guided me.

I would like to convey my sincere appreciation and graditude to my supervisors; Dr. Nevin Darwish, Dr. Magdah Fayek and Dr. Samia Mashali. I would also be very gratuful for Dr. Mahmoud Fakhr El Din for his assitance with my thesis. Thanks for all the time devoted in making constructive comments and suggestions for improving this thesis. 
I thank my family and friends for thier moral and support. I specifically thank, dina said, noha mansour and marwah shafee. Thanks, for being always there when I needed you the most. I hope to express my gratiude to all of you for supproting me and praying for me through all these years. 
 %Finally, I am very grateful to my dear parents, my family, and my friends whom I consider
%as my sisters. I would like to thank particularly Radwa Aboudina, Fatma Nada, Maha
%Nabil, Haidi Badr, and Marwa Kamal. Thank you all for being always there when I needed
%you most. Thank you for believing in me and supporting me through all these years. I think
%without your support and your prayers, none of this work would be accomplished.
%ii
 \end{acknowledgments}                                   %

                                                        % FOREWORD PAGE
%\begin{foreword}                                        %======================
    											                             % Delete these if
%\end{foreword}                                          % no foreword page.

                                                        % OTHER PAGES
                                                        %======================
\tableofcontents                                        % Always needed...
%\listoftables                                           % Delete if no tables.
\listoffigures                                          % Delete if no figures.

%------------------------------------------------------------------------------
%  Document body - Place text into individual include files, such as
%                  as "chap1.tex", or replace each "\include{}" statement
%                  with your actual document text.
%------------------------------------------------------------------------------
\begin{abstractpage}          
% ABSTRACT PAGE
%==============================
% Creates the abstract page.
%   Your text goes here. Just
%   follow the rules given in
%   the LaTeX manual on how to
%   enter/format text using
%   LaTeX commands.

% didn't finish it all Need some addint
The need for new level of human computer interaction had lead the researchers to attempt improve the current interfaces. Even though the area of research progressed quickly in speech and handwriting systems, gesture and sketch understanding systems are still starting to get noticed.

Sketch recognition is defined as the process of identifying the symbols the user draws using a single or multiple strokes. Users draw strokes using a pen and the system immediately interprets their strokes into objects that can be easily manipulated. This research uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. Two algorithm are used to divide strokes, the first algorithm \textit{ALgS1} divides stroke to polygons. The second one \textit{AlgS2}, divides stroke to segments of either lines or curves. The research focuses on the effect of PSO segmentations to the sketch recognition systems. The final recognition is achieved using a set of geometrical and global shape properties features. A SVM (support vector machine) classifier is used to correctly identify shapes.

Experiments were conducted on three different datasets; Hs-DB a benchmark dataset which contains simple presentation symbols, a LD-DB a set of logic design symbols and EL-DB Electrical symbols. Results show that using \textit{ALgS2} to segment stroke improves segmentation than known algorithms. \textit{ALgS1} achieved good performance on EL-DB and Hs-DB but performed poorly on LD-DB due to high number of curves in every symbol. %The segmentation improvement is tested by final symbol recognition accuracy.

\end{abstractpage}                              % This ends the page.

                                                % CHAPTERS
                                                %==============================
\include{ChapterIntroduction}                              % The best way to organize your
\include{chapterSurvey}                              % document is probably to keep
\include{System}
\include{Results}
\include{ConFut}
                                                % in its own file.  
%\include {Mybibliography}
\bibliography{../../neededfiles/Bibliographies/Mybibliography}     
                                                % THE APPENDIX
                                                %==============================
%\appendix                                       % Include appendix files as
  %\begin{singlespace}                           % needed or delete this block
  %\include{appendix_a}                          % of lines if no appendix.
  %\include{appendix_b}                          %
 % \include{appendix_c}                          %
%\end{singlespace}                               %

\end{document}                                  % Ends the entire document...