### 2009-05-31 trunk/bib/old bib/bib.tex
initall add

@article{ SketchRec1, 
  title ="  Sketch Recongition for software engineering  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Levet Burak  Kara   and  ",
}
@article{ sketchint2, 
  title ="  sketch interfaces: Towards more human interface design  ", 
 YEAR = " 2001  ",
Month= "5",
 author =" James Landay   and  Brad A. Myers   and  ",
}
@article{ Improving 3, 
  title ="  Improving sketching by utilizing Haptic feedback  ", 
 YEAR = " 2002  ",
Month= "5",
 author =" Chris Raymaekers   and  Gert Vansichem   and  Frank van Reeth   and  ",
}
@article{ early proc4, 
  title ="  early processing for sketch understanding  ", 
 YEAR = " 2001  ",
Month= "12",
 author =" Tevfik Metin sezgin   and  Thomas Stahovich   and  Randall Davis   and  ",
}
@article{ recognizin5, 
  title ="  recognizing multi stroke symbols  ", 
 YEAR = " 2002  ",
Month= "12",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  Tolga kurtoglu    and  Chris calhoun   and  ",
}
@article{ interpreti6, 
  title ="  interpreting schematic sketches using physical reasoning  ", 
 YEAR = " 2002  ",
Month= "12",
 author =" Thomas Stahovich   and  Tolga kurtoglu    and  ",
}
@article{ a sketch b8, 
  title ="  a sketch based interface for the design and analysis of a simple vibratory mechanical system  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  Leslie Gennari   and  ",
}
@article{ Experience9, 
  title ="  Experiences developing and E-whiteboard based circuit designer  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Ray Liu   and  Lisa  Wong    and  John Grundy   and  ",
}
@article{ recognitio10, 
  title ="  recognition of freehand sketch using mean shift  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" bo yu   and  ",
}
@article{ a frame wo11, 
  title ="  a frame work for multi domain sketch recognition  ", 
 YEAR = " 2002  ",
Month= "5",
 author =" Christine Alverado   and  Randall Davis   and  Michael oltmans   and  ",
}
@article{ Specifying12, 
  title ="  Specifying Gesture by example  ", 
 YEAR = " 1991  ",
Month= "4",
 author =" Dean Rubine    and  ",
}
@article{ Hand drawn13, 
  title ="  Hand drawn  maps for robot navigation  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Majorie Skubic    and   Pascal matsakis   and  Sam bilsard   and  ",
}
@article{ Tahuti : a14, 
  title ="  Tahuti : a geometrical Sketch Recognition  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Tracy hammond   and  Randall Davis   and  ",
}
@article{ Functions 15, 
  title ="  Functions in design idea generation meetings  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Remko van der Lugt   and  ",
}
@article{ Toward com16, 
  title ="  Toward computational model of sketching   ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Kenneth D. Forbus   and  Jeffery M.  Usher   and  Ronald W. Ferguson   and  ",
}
@article{ A domain  17, 
  title ="  A domain  independent system for sketch recognition   ", 
 YEAR = " 2001  ",
Month= "4",
 author =" bo yu   and  Shijie Cai   and  ",
}
@article{ A cognitiv18, 
  title ="  A cognitive approach to sketch understanding  ", 
 author =" Kenneth D. Forbus   and  Ronald W. Ferguson   and  ",
}
@article{ Sketching 19, 
  title ="  Sketching for knowledge capture a progress report  ", 
 author =" Kenneth D. Forbus   and  Jeffery M.  Usher   and  ",
}
@article{ A Sketch i20, 
  title ="  A Sketch interface for mobile robots  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Majorie Skubic    and  Craig Bailey   and  George Chornis   and  ",
}
@article{ Robust Ske21, 
  title ="  Robust Sketched symbol fragmentation using templates  ", 
 YEAR = " 2004  ",
Month= "12",
 author =" Heloise Hse   and  Michael Shilman   and  A. Richard Newton   and  ",
}
@article{ Computer A22, 
  title ="  Computer Aided sketching to capture preliminary design  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" Beryl Plimmer   and  Mark Apperley   and  ",
}
@article{ handling o24, 
  title ="  handling over traced strokes in hand drawn sketches  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Tevfik Metin sezgin   and  Randall Davis   and  ",
}
@article{ Fluid Sket25, 
  title ="  Fluid Sketches : continuous Recognition and morphing of simple hand drawn shapes  ", 
 YEAR = " 2000  ",
Month= "4",
 author =" James Avro    and  Kevin Novins   and  ",
}
@article{ statical v26, 
  title ="  statical visual language models for ink parsing.  ", 
 YEAR = " 2000  ",
Month= "4",
 author =" Michael Shilman   and  A. Richard Newton   and  Hannna Psula    and  Stuart Russell   and  ",
}
@article{ speed and 27, 
  title ="  speed and sketching for multimodal design  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Randall Davis   and  Aron Adler   and  ",
}
@article{ Generic an28, 
  title ="  Generic and HMM based approaches to freehand sketch recognition  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Tevfik Metin sezgin   and  ",
}
@article{ SketchRead29, 
  title ="  SketchRead : A Multi domain sketch recognition engine  ", 
 YEAR = " 2004  ",
Month= "4",
 author =" Christine Alverado   and  Randall Davis   and  ",
}
@article{ LADDER: a 30, 
  title ="  LADDER: a language to describe drawing display and editing in sketch recognition  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Tracy hammond   and  Randall Davis   and  ",
}
@article{ hierarchic7, 
  title ="  hierarchical parsing and recognition of hand sketch diagrams  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  ",
}
@article{ A domain d31, 
  title ="  A domain description language for sketch recognition  ", 
 YEAR = " 2003  ",
Month= "4",
 author =" ",
}
@article{ Three main23, 
  title ="  Three main concerns is sketch recognition an and approach to address them  ", 
 YEAR = " 2002  ",
Month= "4",
 author =" James V. Mahoney    and  Markus p. j. formhez   and  ",
}
@article{ learning s1, 
  title ="  learning structural descriptions form examples   ", 
 YEAR = " 2005  ",
Month= "8",
 author =" ",
}
@article{ Interactio45, 
  title ="  Interactional Context in Sketch Understanding  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Patrick G.T. Healey   and  Simon Garrod   and  Nick Fay   and  John Lee   and  Jon Oberlander   and  ",
}
@article{ Polygon Re31, 
  title ="  Polygon Recognition in Sketch- Based interfaces with immediate and continous feedback  ", 
 YEAR = " 2003  ",
Month= "11",
 author =" Peter Agar   and  Kevin Novins   and  ",
}
@article{ A Multi-Le32, 
  title ="  A Multi-Level Sketching Tool for pencil_and_paper animation   ", 
 YEAR = " 2002  ",
Month= "11",
 author =" Frank van Reeth   and  Fabian Di Fiore   and  ",
}
@article{ Mixing Ima33, 
  title ="  Mixing Images and Sketches for Retrieving Vector Drawings  ", 
 YEAR = " 2004  ",
Month= "11",
 author =" Alfredo Ferreira   and  Manuel J. Fonseca   and  Joaquim A. Jorge   and  Mario Ramalho   and  ",
}
@article{ Analysis o34, 
  title ="  Analysis of Engineering Drawing : State of Art and Challenges  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Karl Tombre   and  ",
}
@article{ sketch und36, 
  title ="  sketch understanding in design: overview of work at the mit ai lab  ", 
 YEAR = " 2002  ",
Month= "12",
 author =" Randall Davis   and  ",
}
@article{ Sketching 37, 
  title ="  Sketching for military courses of action diagrams  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Kenneth D. Forbus   and  Jeffery M.  Usher   and  Vernell Chapman   and  ",
}
@article{ functional38, 
  title ="  functional and formal reasonoing in architectural sketches  ", 
 YEAR = " 2005  ",
Month= "7",
 author =" Ellen Yi- Luen Do   and  ",
}
@article{ efficient 39, 
  title ="  efficient search space exploration for sketch recognition  ", 
 YEAR = " 2004  ",
Month= "9",
 author =" Tevfik Metin sezgin   and  Randall Davis   and  ",
}
@article{ Drawing ma40, 
  title ="  Drawing marks, acts, and reacts: toward a computational sketching interface for architectural design  ", 
 YEAR = " 2005  ",
Month= "7",
 author =" Ellen Yi- Luen Do   and  ",
}
@article{ incrementa41, 
  title ="  incremental sketch understanding for intention extraction in sketch-based user interfaces  ", 
 YEAR = " 2003  ",
Month= "10",
 author =" Yang Li    and  ",
}
@article{ preserving42, 
  title ="  preserving the freedom of paper in a computer based sketch tool  ", 
 YEAR = " 2001  ",
Month= "12",
 author =" Christine Alverado   and  Randall Davis   and  ",
}
@article{ sketching 43, 
  title ="  sketching user interfaces with visual patterns  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Manuel J. Fonseca   and  Joaquim A. Jorge   and  Anabela Caetano   and  Neri Goulart   and  ",
}
@article{ stretch-a-44, 
  title ="  stretch-a-sketch: a dynamic diagrammer  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Mark D Gross   and  ",
}
@article{ mathpad2 :46, 
  title ="  mathpad2 : a system for the creation and exploration of mathematical sketchs  ", 
 YEAR = " 2005  ",
Month= "8",
 author =" Joseph J. La Viola Jr.   and  Robert C. Zelezink   and  ",
}
@article{ agent base47, 
  title ="  agent based sketch recognition  ", 
 YEAR = " 2003  ",
Month= "8",
 author =" Graham Mackenzie   and  ",
}
@article{ an image-b48, 
  title ="  an image-based trainable symbol recognizer for sketch-base interfaces  ", 
 YEAR = " 2004  ",
Month= "9",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  ",
}
@article{ combining 49, 
  title ="  combining geometry and domain knolwdge to interpret hand drawn diagrams  ", 
 YEAR = " 2004  ",
Month= "12",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  Leslie Gennari   and  ",
}
@article{ robust rec50, 
  title ="  robust recognigion and interpretation of mechanical design sketches  ", 
 YEAR = " 1999  ",
Month= "10",
 author =" Christine Alverado   and  ",
}
@article{ sim-u-sket51, 
  title ="  sim-u-sketch: a sketch-based interface for simulink  ", 
 YEAR = " 2005  ",
Month= "7",
 author =" Levet Burak  Kara   and  Thomas Stahovich   and  ",
}
@article{ smart sket56, 
  title ="  smart sketchpad-an on line  graphics recognition system  ", 
 YEAR = " 2001  ",
Month= "11",
 author =" Liu Wenyin   and  Wenjie Qian   and  Rong Xiao   and  Xiangyu jin   and  ",
}
@article{ on line gr35, 
  title ="  on line graphics recognition state of the art   ", 
 YEAR = " 2003  ",
Month= "12",
 author =" Liu Wenyin   and  ",
}
@article{ etcha sket52, 
  title ="  etcha sketches : lessons learned from collecting sketch data  ", 
 YEAR = " 2004  ",
Month= "8",
 author =" Christine Alverado   and  Randall Davis   and  Michael oltmans   and  ",
}
@article{ sketch rec54, 
  title ="  sketch recognition with continuous feedback based on incremental intention extraction  ", 
 YEAR = " 2005  ",
Month= "1",
 author =" Junfeng Li   and  Xiwen zhang   and  Xiang Ao    and  Guozhong Dai   and  ",
}
@article{ HMM based 53, 
  title ="  HMM based efficient sketch recognition  ", 
 YEAR = " 2005  ",
Month= "1",
 author =" Tevfik Metin sezgin   and  Randall Davis   and  ",
}
@article{ a srg base57, 
  title ="  a srg based online composite graphic recognition strategy for sketch based user interface  ", 
 YEAR = " 2002  ",
Month= "1",
 author =" Rong Xiao   and  Xiangyu jin   and  zheng xing sun   and  bin bin peng   and  ",
}
@article{ Handling A55, 
  title ="  Handling Ambiguity and Errors: visual Languages for calligraphic interation   ", 
 YEAR = " 2001  ",
Month= "10",
 author =" Manuel J. Fonseca   and  Joaquim A. Jorge   and  joao  paulo pereira   and  ",
}
@article{ A Retarget58, 
  title ="  A Retargetable Framework for Interactive Diagram Recognition  ", 
 YEAR = " 2003  ",
Month= "1",
 author =" Edward h. lank   and  ",
}
@article{ A parsing 59, 
  title ="  A parsing technique for sketch recongition  ", 
 YEAR = " 2004  ",
Month= "3",
 author =" gennaro costagliola   and  vincenzo deufemia   and  giuseppe polese   and  michele risi   and  ",
}