### 2011-01-11 trunk/IntroductionLetters/document.tex
Inial import

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\begin{document}

\section{Title}

Hi Frank,
I graduated from the Faculty of Engineering Cairo University, in July 2003. I would like to  investigate the possibility of continuing my PhD studies under your supervision at the University of British Columbia. I obtained my Bsc and Msc in Computer Engineering from  Cairo University, Egypt in  2003 and 2009 respectively.  I worked both in the academic and industrial sectors (mention the position).  I am planning (decided) to pursue a research career by first continuing my Post Graduate studies in a respectful university.
I work at Electronics Research Institute in the 'Signal and Image Processing' research group. My work focuses on conducting research and related activities in the field of Machine learning and Image processing. I also work as a part time Research and Teaching Assistant in American University in Cairo (AUC).  My research in AUC focuses on Arabic Handwriting Recognition especially improving interaction and recognition of Online Arabic handwriting systems.
(put this paragraph first)Even though all my research experience is in Machine learning and Image processing fields, I have vast interest in Software Engineering.   I worked for four years as a Software Engineer for Engnet Company, Egypt. I was responsible for analysis of software requirements and the design of the software systems. After the first year, I was promoted to Team Leader and Senior Software Engineer where I got more involved in software process and how to improve the performance of developers. Since then, Software Engineering became my passion; I spent a lot of time trying to improve (Make it as a task) software development process, helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems.  I read  several  technical papers on software engineering, bug tracking, team collaboration and code analysis systems.  Because of my passion I decided to pursue my studies in PhD in Software Engineering field.
I am currently planning for my PhD studies. I have preference towards studying in Canada where state of the art research exists and particularly in the University of British Columbia due to its international recognition as one of the best Universities in Canada. My research interests are mining code and bug repositories (Be specific about the project name..but I am open to any idea in the field). Also I am interested in code analysis and software team collaboration which I hope falls under your research interests as indicated in the website of University of British Columbia (be specific about the website if it is a group or if its is personal website)

I am wondering about the possibility of pursuing my graduate studies under your  guidance and supervision at University of British Columbia as from the fall of 2011. I write to seek your kind advice.

\end{document}