### 2011-01-11 IntroductionLetters/document.tex
correccting the warp

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\begin{document}

\section{Title}

Hi Frank,

I would like to investigate the possibility of continuing my PhD studies under your supervision at the University of Calgary. My name is Maha and I obtained my Bsc and Msc in Computer Engineering from Cairo University, Egypt in 2003 and 2009 respectively. I worked both in the academic (as a Research Assistant) and industrial (as a Software Engineer) sectors. I am planning to pursue a research career by continuing my Post Graduate studies. My main research interest is in integrating software engineering methods with different human-computer interaction.
Most my research experience is in Machine Learning but I have vast interest in Software Engineering. I worked for four years as a Software Engineer for Engnet, Egypt. After the first year, I was promoted to Team Leader and Senior Software Engineer where I got more involved in software process and how to improve the performance of developers. Since then, Software Engineering be- came my passion; I took charge of improving the software development process, helping developers for better collaborative work and looking for methods that will improve issue and bug tracking systems. I read several technical papers on software engineering, bug tracking, team collaboration and code analysis systems. Because of my passion I decided to pursue my studies in PhD in Software Engineering field. (kindly see my CV for more information.)

Currently, I work at Electronics Research Institute in the 'Signal and Image Processing' research group. My work focuses on conducting research and related activities in the field of Machine learning and Image processing. I also work as a part-time Research and Teaching Assistant in American University in Cairo (AUC). My research in AUC focuses on Arabic Handwriting Recognition especially improving interaction and recognition of Online Arabic handwriting systems.

My research interests are in developing applications to improve interaction of tabletop surfaces and approaches for improving interactions between distributed agile teams. I am also interested in how to help teams to create and maintain useful software but I am open to any idea in the field. I hope these interests falls under your research interests as indicated in the web site of the Agile Software Engineering research group.

I have preference towards studying in Canada where state of the art research exists and particularly in the University of Calgary due to its international recognition as one of the best Universities in Canada. I am wondering about the possibility of pursuing my graduate studies under your guidance and supervision at University of Calgary. I write to seek your kind advice.

\end{document}