### 2011-02-25 MyCv/Resyne.tex

% Generated by GrindEQ Word-to-LaTeX 2010
% ========== UNREGISTERED! ========== Please register! ==========
% LaTeX/AMS-LaTeX

\documentclass{article}

%%% remove comment delimiter ('%') and specify encoding parameter if required,
%%% see TeX documentation for additional info (cp1252-Western,cp1251-Cyrillic)
%\usepackage[cp1252]{inputenc}

%%% remove comment delimiter ('%') and select language if required
%\usepackage[english,spanish]{babel}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[dvips]{graphicx}
%%% remove comment delimiter ('%') and specify parameters if required
%\usepackage[dvips]{graphics}
\linespread{1}

\pdfpagewidth 8.5in
\pdfpageheight 11in

\setlength\topmargin{0in}
\setlength\headheight{0in}
\setlength\headsep{-0.25in}
\setlength\textheight{9.5in}
\setlength\textwidth{6.5in}
\setlength\oddsidemargin{-0.25in}
\setlength\evensidemargin{-0.25in}
\setlength\parindent{0in}
\setlength\parskip{0in}

\title {Curriculum Vitae \\ Maha Mohammed Nabeel El
Meseery\\ \small \textbf{ Address:} $2^{nd}$ Tahreer street, Dokki, Giza 12615,
Egypt. \\ \textbf{Tel:} (+2012)112- 5435, (+202)3336-8292.
\textbf{Email:} melmeseery@aucegypt.edu}
\date{}
\begin{document}

%%% remove comment delimiter ('%') and select language if required
%\selectlanguage{spanish}
\maketitle

 \section*{Education:}
  \begin{description}
\item [2004-2009]  \textbf{ Msc.E., Cairo University, Computer Engineering Department, Giza, Egypt}
  \begin{itemize}
    \item \textit{ Thesis Title:}\textbf{Sketch Recognition And Understanding}.
    \item \textit{Thesis Supervisors:} Prof. Nevin Darwish, Dr Magda Fayek , Prof. Samia
 Mashali , Mahmoud Fakr El Din
 \item \textit{ Graduate Courses:} Graduate Courses: Advanced Neural Networks, Image
 Processing and Computer Vision, Advanced Computer Networks, Advanced Database Systems, Embedded Systems, VLSI Testing, Numerical Analysis, and Technical Writing.

\end{itemize}

 \item[1998-2003]  \textbf{ Bsc.E., Cairo University, Computer Engineering Department Giza, Egypt}
   \begin{itemize}
    \item \textit{Final Grade:}\textbf{Very Good with honors}.
  \item \textit{Graduation project:} Personal Face identification System, \textbf{Distinction.}
  \item \textit{Project Supervisors:} Dr. Madga Fayek.
\end{itemize}
  \end{description}

 \section*{Experience:}
 %\textbf{\underbar{Experience:}}
 \textbf{August 2008- present   [part time]}\\
 \textbf{Research and Teaching Assistant}, American University in Cairo, Egypt,
 Electrical Engineering Department.

 \begin{enumerate}
   \item Conducting and helping develop research in Online Arabic Handwriting
   Recognition. This includes implementing various features extraction and classification methods.
   \item Assisting in collecting, gathering and analyzing Online Arabic
   handwriting Datasets.
%   \item Assisting and developed research in area of Arabic and Latin digit
 %  Recognition.
 \item Implementing different features extraction and classification methods for Arabic and Latin digits and comparing their performance.
 \item Reading, analyzing and reviewing scientific and research papers.
 \item Guiding students in their studies in Lab projects and programming
 assignments.
 \item Conducting tutorials and problem sessions for students in various
 subjects.
\end{enumerate}

 \textbf{January 2007- present}\\
 \textbf{Research Assistant}, Electrical Research Institute, Egypt.
 \begin{enumerate}
\item Working as a member of the Signal Processing Lab in the Computer and
Systems Department.
 \item Assisting in the research project undergoing in the department.
 \item Implementing and developing different algorithms in field of Image
 Processing.
 \item Providing a technical support for other researchers especially in field
 of JavaEE programming.
\end{enumerate}

 \textbf{September 2005- September 2008 [part time from January 2007] }\\
 \textbf{Senior Software Engineer}, EngNet, Egypt.

 \begin{enumerate}

\item Introduced new software procedures in the company to enhance developers'
performance and collaboration.
 \item Analyzed and designed various software projects using UML and software engineering concepts.
 \item Wrote the software requirements and specifications of software
 application.
 \item Supervised and managed the development process for the developers and
 quality engineering teams.
 \item Assisted junior programmers in development using Java, JSP, Struts, RCP and MySQL.
\end{enumerate}

 \textbf{September 2003- September 2005}\\
 \textbf{Teaching Assistant}, Faculty of Computer Science, MSA University, Giza, Egypt.
 \begin{enumerate}
\item Developed lab manuals and leading lab experiments for Software Engineering
Methodologies, Object Oriented Programming, Digital Logic Design, Computer Graphics, Calculus, and Introduction to IT.
 \item Assistant supervisor for senior student's projects (Mainly graduation project using ASP.Net, C\#) .
 \item Developing and delivering short courses in C++, C\#, UML, and ASP.NET.
\end{enumerate}

 \section*{Publications:}
 \begin{enumerate}
  \item Online Library of Scientific Models - A New Way to Teach, Learn, and
Share Learning Experience, Elrefaei Hatem, Aboelfadel Tamer, Elmeseery Maha,
Elmofty Ali, Shafee Marwa, in Proceedings of the International Conference of "Interactive computer aided learning" ICL2007, Villach, Austria.\textbf{\underbar{}}
  \item El Meseery, Maha, Fakhr El Din, Mahmoud., Mashali, Samia., Fayek, Magda.
 \& Darwish, Nevin., Sketch Recognition Using Particle Swarm Algorithms,  IEEE
 International Conference on Image Processing, November,  2009, Cairo, Egypt.
  \item El Meseery, Sketch Recognition and Understanding,  Master Thesis,
  Faculty of Engineering, Computer Engineering Department, Cairo
  University, November, 2009, Egypt.
  \item El Meseery, Maha, Fakhr El Din, Mahmoud, Tarek abouldahab, Sketch
  Recognition using Enhanced Particle Swarm Optimization, Submitted to Pattern
  Recognition Letters, (17 pages Submitted in Dec 2010).
\end{enumerate}
 \section*{Awards:}
  \begin{enumerate}
    \item Jul 2003: Cairo University Award for Academic Distinction, Egypt.
	\item 1999 - 2003: Fellowship, Faculty of Engineering, Cairo University, Giza,
	Egypt.
  \end{enumerate}
 \section*{Computer Skills:}
 \begin{enumerate}
\item Environments: Linux (openSuse) and Windows.
 \item Programming: Java, JSP, Sturts, AJAX, .Net,C/C++, C\#,  ASP.NET, ASP, ADO.Net, ADO, VC6,  XML  Visual Basic, Embedded C, Assembly, Pascal, VHDL, Prolog, Lisp, Matlab, latex.
 \item Database Tools: SQL server 2000, MySql, Oracle and developer 6i.
 \item Analysis and Modeling: UML, Hardware design.
\end{enumerate}

 \section*{Training Courses and Summer jobs:}
 \begin{enumerate}
\item Human Development, Zedny Training Center, Cairo, Egypt (January - March 2005).
 \item MS SQL 2000, MSA University, Giza, Egypt (July 2004).
 \item C\# dot Net and ADO dot Net, MSA University, Giza, Egypt (August 2004).
 \item Web Development (ASP, XML, DOM, DHTML and ADO), Harf Information Technology. Cairo, Egypt (July --September 2002).
 \item DB2, JSP, Java beans, Servelts and IBM Web Sphere, Cairo University, Giza, Egypt (July- August 2001).
 \item Java applets, Training to develop an e-version of the calculus book
 taught to engineering students (August -- September 2000).
 \item MFC and VC6, Cairo University, Giza, Egypt (July 2000).
\end{enumerate}

 \section*{Projects:}

 \begin{enumerate}
\item 2010-Present, Online Handwriting project.
 \item 2009-2010, Arabic Digit Recognition project.
\item 2007-2008, Schedulizier: a project to schedule and manage information of
training companies ( EngNet, Egypt.).
 \item 2006-2007, Scientific SDK: a project to help engineers create scientific
 simulations by automatically generating the user interface for them.( EngNet,
 Egypt)
 \item 2005-2006, Imodelit : a library of scientific Microwave java applets (
 EngNet, Egypt).
 \item 2004, Mouth detection and segmentation using C++ (Master's Project).
 \item 2004, Traffic Control System using micro-controllers (Master's
 Project).
 \item 2003, Using Neural Network to recognize English Characters using Matlab.
 \item 2003, Keyless Entry System using 80468 assembly.
 \item 2003, Implementing a Processor using VHDL.
 \item 2002, Simulation of a single processor operating system using Java.
 \item 2001, Developing a Database for a social security accounting office using Oracle Database.
 \item 2001, Water Filter (Hardware project).
 \item 2000, Snake Game using Pascal.
\end{enumerate}

 \section*{Languages:}

Arabic (Native), English (Fluent), French (poor).

 \section*{Social Activities and Professional memberships:}

\begin{enumerate}
\item  Student member of the Institute of Electrical and Electronics Engineers
(IEEE) since 2002.
\item  Member of the Egyptian Engineering Syndicate (ESE) since 2003.
\item  Member of \textit{Resala} Charitable Society since 1999, responsible for
the following activities:
\begin{enumerate}
\item  Carrying out the big sister role for orphan girls. This includes the following
  \begin{enumerate}
\item Helping developing better education procedures for \textit{Resala}
organization orphans.
\item Attend courses on behaviour modification and  social interaction for
teenagers.
\item Acting as guidance for behavioural, social and mental aspects for orphaned
teenagers.

\item Teach teenagers educational courses in different subjects.
\end{enumerate}
\item  Organizing blood donations campaigns.
\item  Delivering free-charge courses in IT (C, C++, VC6, Windows and office ).
\item  Helping in writing the contents of tutorial for the various courses given at Resala ITcenter.
\end{enumerate}
\item  Member of Egyptian shooting clubs.
\item  Enjoy reading, swimming, cooking, walking and watching movies.
\end{enumerate}
%\newpage
% \section*{Personal Information:}
% \noindent
%  \textbf{ Address: } $2^{nd}$ Tahreer street, Dokki, Giza 12615, Egypt.
%
%  \textbf{ Telephone:} (+2012)112- 5435, (+202)3336-8292.
%
% \textbf{  Email: }  melmeseery@aucegypt.edu

\section*{References:}
%\noindent
\begin{description}

\item[\textbf{Dr. Magda Fayek}] (Professor)

\begin{description}
  \item[Address:]Computer Engineering Dept., Faculty of Engineering, Cairo University, Giza, Egypt.
  \item[Email:] magdafayek@eng.cu.edu.eg.
%\\\textbf{Email: }
\end{description}
\item[\textbf{Dr. Hatem El refaei}] (Associate Professor)

\begin{description}
  \item[Address:]Physics and Mathematics Dept., Faculty of Engineering, Ain Shams University, Cairo, Egypt.
  \item[Email:]hatem@mcit.gov.eg.
\end{description}

\item[\textbf{Dr. Sherif Abdel Azeem}] (Associate Professor)

\begin{description}
  \item[Address:]Electrical Engineering Dept., School of Science and Engineering,American University in Cairo, New Cairo, Cairo, Egypt.
  \item[Email:]  shazeem@aucegypt.edu.
\end{description}

\item[ \textbf{Dr. Mahmoud Fakhr El Deen}] (Researcher)

 \begin{description}
  \item[Address:]Computer and Systems Dept., Electronic Research Institute,
  National Research Center, Tahreer Street, Dokki, Cairo, Egypt.
  \item[Email:]mafakhr@mcit.gov.eg.
\end{description}
\end{description}
\end{document}

% == UNREGISTERED! == GrindEQ Word-to-LaTeX 2010 ==