### 2011-08-11 FacePaper/aiml.tex
adding all the files for first time

% Template AIML-ICGST-LaTex-Template
% 1/06/04

\documentclass[a4paper, twoside]{article}

\usepackage{icgst}
\begin{document}
\input{files/head}
\input{files/abstract}
\input{files/introduction}
\input{files/theory}
\input{files/conclusion}
\input{files/ack}

\bibliographystyle{plain}
\bibliography{files/icgst}

\newpage
\section*{Biographies}

\parpic{\includegraphics[height=3cm]{pics/guest}} \textbf{Author One} \small Author One received his B.Sc. from .......  at ....... and his M.Sc. from ........ at ....... Currently he is working for ....... and he is a PhD student at University of ....... in the ........ Laboratory under the supervision of ......... His research interests include .........,  .......... and ...........\\

\parpic{\includegraphics[height=3cm]{pics/guest}}
\textbf{Author Two} \small is currently Professor at University of ....., in the ....... Laboratory. He is Senior Member of the ...... He is Editor Associate of the ....... on ....... He received ........ award in ...... He is doing his research work in ...... for ........ His main research interests are ......, ....., .... and .......\\

\end{document}