### 2011-08-11 FacePaper/files/abstract.tex

% AIML ICGST LaTeX Template

\begin{abstract}
%% Text of abstract
This paper introduce a system that uses Particle Swarm Optimization

\end{abstract}

%\textit{\textbf{Keywords:} Autonomous robot, Stochastic control, Kalman filter,
%Fuzzy logic, Neural Network, Adaptive navigation.}