### 2011-08-11 FacePaper/files/conclusion.tex

% AIML ICGST LaTeX Template 
\section{Conclusion}
\label{Sec:Con}
This is the conclusion. The major contributions of this article arise from the formulation of a new approach, spectral analysis, to the modelling and identification of XYZ and ABC features that provide improved computational efficiency in the positioning techniques. By manipulating the manner in which feature information of both A and BVN signatures is incorporated into the model, it can be shown that significant improvements in the performance of the algorithm can be realised. Moreover, the simplicity and the efficiency of dynamic pose tracking techniques succeeded to improve the robot pose estimation process.