### 2011-08-11 FacePaper/files/introduction.tex

% AIML ICGST LaTeX Template 
\section{Introduction}
\label{Sec:int}
We consider that it is necessary to decide a state of the network load in order to adjust an image information quantity responding a state of the network.  In this paper, A state of the network load is decided by using a ratio of send and receive of image data communicated between slave site and master site, a receive interval of image receive report and a send completion time of \textbf{image data}.  The master site reports a received image data quantity to the slave site every a certain time.  The slave site decides a state of the network load by making a comparison between received report and image data quantity transmitted until it receive reports.  Simultaneously, the slave site decides that the network load is large, in case that it can't receive the report inside a certain time.  Also the slave site measures a transmission completion time of image data.\\

To design Service Robots for elderly and disabled people is a very demanding task. There are fundamental differences in the necessities, compared with industrial robot applications, as well as to several other service robot applications. The tasks vary in wide areas, depending on the actual situation of the user and they cannot be pre programmed as in other robotic applications. While all service robots must be designed for operation in clustered, a priory unknown and variable environments, new demands arise for the support of the target group. The interaction with the user and the variety of the supporting tasks may be mentioned here. On the other hand no total autonomy is demanded. It is rather important to use the cognitive abilities of the users and support compensate they only in areas where physical or cognitive impairments exist. Another limitation is the number of units which may be sold. Closely related to the small number compared with other industrial products is the price of the today available components. The design of a robotic aid must consider these limitation but also possibilities which arise due to the specific users. For the near future, the primary target group are disabled people with serious physically handicaps but only small mental disabilities if at all. Opposite to other service robot applications the cognitive abilities can and must be used to support and simplify the controller design, esp. for environment identification and scene interpretation. The remaining automation of the tasks is still very demanding, but the sensor data interpretation is eased if the user support the automation system. The paper will discuss the design philosophy which results from this approach. An electric wheelchair with an added robotic arm serves as an example. The main target group of this unit are paraplegics ore people with similar handicaps. \\

The remainder of the paper is organized as follows: Section (\ref{Sec:Th}) focuses on ....... Section (\ref{Sec:Con}) emphasizes on ...........